Lesson 5
========================================================

### Multivariate Data
Notes:

### Loading dataset

```{r}
setwd("F:/Datasets")
pf <- read.delim("pseudo_facebook.tsv")
```

```{r}
str(pf)
```

### Loading required libraries

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
```


***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
```

```{r}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = "summary", fun.y = median)
```

```{r}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age,gender) %>%
  summarise(mean_friend_count = mean(friend_count), median_friend_count = median(friend_count), n = n()) %>%
  arrange(age,gender)
```

```{r}
head(pf.fc_by_age_gender)
```


***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
pf.fc_by_age_gender %>% ggplot(aes(x = age, y = median_friend_count)) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>%
  filter(!is.na(gender)) %>%
  select(age, gender, median_friend_count) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)
```

```{r}
pf.fc_by_age_gender.wide <- tbl_df(pf.fc_by_age_gender.wide)
pf.fc_by_age_gender.wide
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.wide %>% ggplot(aes(x = age, y = female / male)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

```{r}
str(pf$year_joined)
```


***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
```

```{r}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009, 2011, 2012, 2014))
```


```{r}
table(pf$year_joined.bucket)
```



***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
subset(pf, !is.na(gender) & !is.na(year_joined.bucket)) %>%
  ggplot(aes(x = age, y = friend_count)) + 
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
subset(pf, !is.na(gender) & !is.na(year_joined.bucket)) %>%
  ggplot(aes(x = age, y = friend_count)) + 
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205 

What is the maximum friend rate? 417.0000

```{r Friendships Initiated}
subset(pf, pf$tenure >= 1) %>%
  ggplot(aes(x = tenure, y = friendships_initiated / tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

```{r}
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket), stat = "smooth")

```


***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:


```{r}
setwd("F:/Datasets")
yo <- read.csv("yogurt.csv")
str(yo)
```

```{r}
summary(yo)
```

```{r}
yo$id <- as.factor(yo$id)
```

```{r Histograms Revisited}
ggplot(aes(x = price), data = yo) +
  geom_histogram(fill = I("#F79420"), binwidth = 10)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
table(yo$price)
```

```{r}
yo$all.purchases <- with(yo, strawberry + plain + mixed.berry + blueberry + pina.colada)
str(yo)
```


```{r}
yo %>%
  ggplot(aes(x = all.purchases)) +
  geom_histogram(fill = I("#fe7722"), color = "black", binwidth = 1)
```


***

### Prices over Time
Notes:

```{r Prices over Time}
yo %>% 
  ggplot(aes(x = time, y = price)) + 
  geom_point(alpha = 1/5, position = position_jitter(h = 0))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(420)
sample.ids <- sample(levels(yo$id), 16)
```

```{r}
subset(yo, id %in% sample.ids) %>%
  ggplot(aes(x = time, y = price)) +
  facet_wrap(~id) + 
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)

```


***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r}
library(GGally)
theme_set(theme_minimal(20))
```


```{r}
set.seed(1836)
pf_subset <- pf[,c(2:15)]
names(pf_subset)
```

```{r}
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```

```{r}
set.seed(1836)
data_pf <- pf[sample.int(nrow(pf),1000),]
with(data_pf, cor.test(friendships_initiated, friend_count))
with(data_pf, cor.test(age,mobile_likes))
```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
setwd("F:/Datasets")
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

